---
title: "Assignment 3"
author: "Alicia Canales and Zoe Zhou"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    code_folding: 'hide'
---

```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(here)
library(patchwork)
library(ggplot2)
```

### Profit model
```{r}
## Sourcing compute crop yield fucntion and compute profit function
source('../functions/compute_crop_yield.R')
source('../functions/compute_profit.R')
set.seed(123)

## reading in data
climate_data <- read.table(here('data', 'clim.txt'), header = TRUE)

## running profit model on climate data
almond_profit <- compute_profit(climate_data)
```

### Informal Sensitivity Model for Cost
```{r}
## sampling different parameter values for cost variable
cost_var <- rnorm(mean = 20, sd = 10, n = 20)

## mapping the compute profit function over the new cost variables
cost_sensitivity <- map(cost_var, ~compute_profit(climate_data, cost = cost_var))
head(cost_sensitivity)

## extracting values into a dataframe
cost_sen <- map_df(cost_sensitivity,`[`, c('year', 'profit')) 
```

### Informal Sensitivity Model for Labor
```{r}
## sampling different parameter values for labor variable
labor <- rnorm(mean = 50, sd = 15, n = 20)  

## mapping the compute profit function over the new labor variables
labor_sensitivity <- map(labor, ~compute_profit(climate_data, labor = labor))
head(labor_sensitivity)

## extracting values into a dataframe
labor_sen <- map_df(labor_sensitivity,`[`, c('year', 'profit')) 
```

### Sensitivity Results
```{r}
## Cost Plot
p1 <- ggplot(cost_sen, aes(x = year, y = profit, group = year)) +
  geom_boxplot(fill = 'blue') +
  labs(x = 'Year', y = 'Total Profit ($)') +
  ggtitle('Sensitivity Analysis of Cost') +
  theme_minimal() 


## Labor Plot
p2 <- ggplot(labor_sen, aes(year, profit, group = year)) +
  geom_boxplot(fill = 'orange') +
  labs(x = 'Year', y = 'Total Profit ($)') +
  ggtitle('Sensitivity Analysis of Labor') +
  theme_minimal()

combined <- p1 / p2
combined
```

*Interpretation of Model*

This sensitivity analysis focused on the profit function, highlighting how inputs like cost and labor directly impact profitability. It identified labor cost as the main factor influencing profit fluctuations, demonstrated by boxplots showing greater variance for labor costs. In the profit function, labor and cost are multiplied to compute total costs, subtracted from income to determine net profit. Thus theoretically, changes in labor or cost should equally affect profit, given their proportional relationship in the calculation. Discrepancies between two sensitivity analyses may arise from scaling differences.
