---
title: "Assignment 6"
author: "Alicia Canales"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    code_folding: 'hide'
---

```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(ggplot2)
library(deSolve)
library(sensitivity)
```

### ODE Solver
```{r}
source('../functions/compute_forest_growth.R')

# initial conditions over 300 years
years <- seq(1,300,1)

# initial forest size 
initial_c <- 10

# initial forest parameters
params <- list(k = 250, r = 0.01, g = 2, threshold = 50)

# ode solver with initial conditions 
ode_results <- as.data.frame(ode(initial_c, years, compute_forest_growth, params))

# changing names of columns to year and carbon (forest size)
colnames(ode_results) = c('year', 'carbon')

# plot of ODE results 
ggplot(ode_results, aes(x = year, y = carbon)) +
  geom_point(alpha = .5, color = 'blue') +
  labs(x = 'Year', y = 'Forest Size (Carbon)', title = 'ODE Solver Results') +
  theme_minimal()
```

### Sobol Global
```{r}
# still using the same initial forest size

# using two sets of parameters for sensitivity analysis
# first set of parameters
np=2000
r = rnorm(mean = 0.01, sd = 0.01/10, n = np)
g = rnorm(mean = 2, sd = 2/10, n = np)
k = rnorm(mean = 250, sd = 250/10, n = np)
thres = rnorm(mean = 50, sd = 50/10, n = np)

X1 = cbind.data.frame(r = r, g = g, k = k, threshold = thres)

# second set of parameters
r = rnorm(mean = 0.01, sd = 0.01/10, n = np)
g = rnorm(mean = 2, sd = 2/10, n = np)
k = rnorm(mean = 250, sd = 250/10, n = np)
thres = rnorm(mean = 50, sd = 50/10, n = np)

X2 = cbind.data.frame(r = r, g = g, k = k, threshold = thres)

# fixing negative values
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# sobol object
sens_P = sobolSalt(model = NULL,X1, X2, nboot = 300)

# renaming columns
colnames(sens_P$X) = c('r','g','k', 'threshold')

# checking the parameter sets
head(sens_P$X)
 
# putting all the parameters into one list to run it through ODE
sobol_parms = list(r = sens_P$X[1,'r'], g = sens_P$X[1,'g'],k = sens_P$X[1,'k'], threshold = sens_P$X[1,'threshold'])

# ode solver with sobol params using just the first set
sobol_ode_result = ode(initial_c, years, compute_forest_growth, sobol_parms)

# turning into data frame
result <- as.data.frame(sobol_ode_result)

# renaming column names 
colnames(result) = c('year', 'carbon')
```

### Running ODE with Sobol on all parameters using wrapper functions
```{r}
# computing metrics function to output the max carbon size and what year it occurs in 
compute_forest_metrics <- function(result){

  # max carbon size 
  max = result[result$carbon == max(result$carbon), ]
  
  # what year the max carbon size is in to check that its coming from year 300 
  year = max$year
  
  # the value of carbon in that year
  forest_size = max$carbon
  
  return(list(max_forest_size = forest_size, year = year))
}

# checking to see if it runs with one parameter set
test = compute_forest_metrics(result)

# creating wrapper function that will run everything

wrapper <- function(r, g, k, threshold, initial_c, time, odefunction, metricfunction){

  # creating initial parameter list with carrying capacity, growth rates (r, g), and threshold
  params <- list(k = k, r = r, g = g, threshold = threshold)
  
  # running ode solver
  ode_results <- ode(initial_c, years, compute_forest_growth, params)

  # changing names of columns to year and carbon (forest size)
  colnames(ode_results) = c('year', 'carbon')
  
  # getting the metrics from ode solver results
  metrics = compute_forest_metrics(as.data.frame(ode_results))
  
  return(metrics)
}

# testing on one set to make sure it runs
wrapper(r = 0.01, g = 0.02, k = 150, threshold = 50, initial_c = 3, time = seq(1, 100),
         odefunction = compute_forest_growth, metricfunction = compute_forest_metric)
```

### Running wrapper on all parameters
```{r}
allresults = as.data.frame(sens_P$X) %>% 
  pmap(wrapper, initial_c = initial_c, time = years, odefunc = compute_forest_growth, metricfunction = compute_forest_metric)

# extract out results from pmap into a data frame
allres = allresults %>% 
  map_dfr(`[`,c('max_forest_size','year'))
```

### Running Sensitivity analysis
```{r}
# Running sensitivity analysis on the parameters that generated the max carbon
sens_P_max_carbon = sensitivity::tell(sens_P, allres$max_forest_size)

# changing the name of the columns
rownames(sens_P_max_carbon$S) = c('r', 'g', 'k', 'threshold')

# first order sobol indices
sens_P_max_carbon$S # 0.34807077, 0.25005313, 0.37849225, 0.07054889

# total sobol indices
rownames(sens_P_max_carbon$T) = c('r', 'g', 'k', 'threshold')
sens_P_max_carbon$T # 0.33878815, 0.22275324,	0.37344403, 0.06253561	

# combining the results and the parameters into one dataframe
parms <- as.data.frame(sens_P_max_carbon$X)
both <- cbind.data.frame(parms, max_carbon = sens_P_max_carbon$y)

# making it tidy to make a boxplot
tmp = both %>% 
  pivot_longer(cols = 1:4,names_to="metric", values_to="value")

# boxplot of maximum forest size
ggplot(both, aes(y = max_carbon))+
  geom_boxplot(color = 'green4') +
  labs(y = 'Maximum Forest Size (C)', title = 'Forest Size') +
  theme_minimal() +
  guides(color = guide_legend(title = '')) +
  theme(axis.text = element_text(size = 12),
        legend.text = element_text(size = 11))


## Just Extra
# boxplot using the parameters that generated the max forest size
ggplot(tmp, aes(x = metric, y = value, color = metric))+
  geom_boxplot() +
  labs(x = 'Metric', y = 'Value', title = 'Sensitivity of 4 Parameters') +
  theme_minimal() +
  guides(color = guide_legend(title = '')) +
  theme(axis.text = element_text(size = 12),
        legend.text = element_text(size = 11))
```

5. Sobol Indices
*S: r = 0.34807077, g =  0.25005313, k = 0.37849225, threshold = 0.07054889* and 
*T: r = 0.33878815, g = 0.22275324,	k = 0.37344403, threshold = 0.06253561*	

6. The sobol indices tell us that the parameter r and k have the highest influence on the forest size since they have the highest values associated with them. This makes sense because the growth rate and the capacity of the forest will directly determine the units of carbon that can be sequestered. We see this in the first boxplot where the forest size is variable depending on the rate at which the forest is growing and the capacity it can get to. When thinking about this in regards to climate change, both of these parameters will be vulnerable to deforestation, soil quality, droughts, and flooding influencing to forests' ability to bounce back or not. In the boxplot, we see that the carrying capacity (k) producing the highest carbon value is the most variable which could mean that in 300 years the size of the forest is going to decrease. 

